\bf Hierarchical Network Structure Promotes Dynamical Robustness

### 2015-07-22 tex/introduction.tex
update introduction

Rather than evaluating system stability for a particular model of a biological network over physiological timescales, we are interested in evaluating dynamical robustness over evolutionary timescales where the form of the most accurate underlying model is itself subject to change. Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e. parameters, structure of the rate functions, etc.), and these changes may occur at any level of the hierarchy including metabolic, gene-regulatory, population, and ecosystem. It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states and even to the geometry of their state spaces. However, what must remain invariant on such evolutionary timescales is the higher-level property that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process subject to environmental constraints. This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon1962,Simon2002}. What is important over at least moderate evolutionary timescales is then the conditional probability, $R$, given a system is in a stable state, that upon modifications to its structure in the context of environmental fluctuations, it remains stable, regardless of where the stable state is located within the state space \ref{fig:robustnessconcept}. For the purpose of this investigation then, we quantify dynamical robustness in this way.
% as the probability of stability to potentially large perturbations of the model specifying the system dynamics for systems which have already been determined to be linearly stable.

>>>>

Rather than evaluating system stability for a particular model of a biological network over physiological timescales, we are interested in evaluating dynamical robustness over evolutionary timescales where the form of the most accurate underlying model is itself subject to change. Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e. parameters, structure of the rate functions, etc.), and these changes may occur at any level of the hierarchy including metabolic, gene-regulatory, population, and ecosystem. It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states and even to the geometry of their state spaces. However, what must remain invariant on such evolutionary timescales is the higher-level property that the system possess a relatively high overall probability of remaining near a stable state upon modifications to the underlying dynamical process subject to environmental constraints. This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon1962,Simon2002}. What is important over at least moderate evolutionary timescales is then the conditional probability, $R$, given a system is in a stable state, that upon modifications to its structure in the context of environmental fluctuations, it remains stable, regardless of where the stable state is located within the state space \ref{fig:robustnessconcept}. For the purpose of this investigation then, we quantify dynamical robustness in this way.
% as the probability of stability to potentially large perturbations of the model specifying the system dynamics for systems which have already been determined to be linearly stable.
